Inferring the Recombinant Genealogy of a Pandemic Pathogen in Real Time

### 2025-09-12 latex/main.tex
Updates from Overleaf

% See https://github.com/jeromekelleher/sc2ts-paper/issues/628
% Referenced by the main Results.
The ARG captures important deletion events (Table~\ref{tab:major_dels}) and highly recurrent deletion events (Table~\ref{tab:rec_dels}).

Li et al.\cite{Li2023indels} performed a large-scale analysis of indels
using a global GISAID dataset (sampled up to August, 2022).
We examined the major indels reported in Table S1 of Li et al.\cite{Li2023indels}
which have a minimum frequency of 1\% (count of sequences divided by 9,149,680 sequences analyzed).
All of these indels are deletions and have high frequencies in the ARG
(sequence count divided by the total number of samples),
except a 9-base deletion at 11283, a 9-based deletion at 21987, and a 9-base deletion at 686.

\begin{table}
\centering
\caption{Major deletion events in the ARG. These events are inherited by at least 10,000 samples in the ARG. Count is the number of samples inheriting a deletion. The Pango lineage assigned to the node above which an event occurs is shown. Abbreviations: NSP, non-structural protein; ORF, open reading frame; S, Spike; S1-NTD, N-terminal domain in Subunit 1.}
\label{tab:major_dels}
\begin{tabular}{rrcrcrr}
\toprule
Start & Length & Region & Node & Pango lineage & Count & Frequency (\%)\\
\midrule
28271 & 1 & non-coding & 1436808 & B.1 & 1118363 & 45.0561 \\
28248 & 6 & ORF8 & 220186 & B.1.617.2 (Delta) & 1114669 & 44.9073 \\
22029 & 6 & S / S1-NTD & 200039 & B.1.617.2 (Delta) & 1112504 & 44.8200 \\
11288 & 4 & ORF1ab / NSP6 & 1436802 & B.1.1.529 & 871782 & 35.1220 \\
28362 & 9 & N & 1436802 & B.1.1.529 & 871774 & 35.1216 \\
21633 & 9 & S / S1-NTD & 822854 & BA.2 (Omicron) & 532072 & 21.4359 \\
11292 & 5 & ORF1ab / NSP6 & 822854 & BA.2 (Omicron) & 531894 & 21.4287 \\
6513 & 3 & ORF1ab / NSP3 & 851246 & BA.1 (Omicron) & 340757 & 13.7283 \\
11283 & 5 & ORF1ab / NSP6 & 851246 & BA.1 (Omicron) & 339867 & 13.6924 \\
21765 & 6 & S / S1-NTD & 851246 & BA.1 (Omicron) & 339681 & 13.6849 \\
21987 & 9 & S / S1-NTD & 851246 & BA.1 (Omicron) & 339651 & 13.6837 \\
22194 & 3 & S / S1-NTD & 851246 & BA.1 (Omicron) & 312706 & 12.5982 \\
28271 & 1 & non-coding & 86456 & B.1.1.7 (Alpha) & 290446 & 11.7014 \\
11288 & 9 & ORF1ab / NSP6 & 86456 & B.1.1.7 (Alpha) & 290408 & 11.6998 \\
21991 & 3 & S / S1-NTD & 86456 & B.1.1.7 (Alpha) & 288831 & 11.6363 \\
21765 & 6 & S / S1-NTD & 86456 & B.1.1.7 (Alpha) & 290438 & 11.7010 \\
21765 & 6 & S / S1-NTD & 1265302 & BA.4 (Omicron) & 190425 & 7.6718 \\
686 & 9 & ORF1ab / NSP1 & 2698748 & BA.4 (Omicron) & 23330 & 0.9399 \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}
\centering
\caption{Highly recurrent deletions. These are the top 20 deletions in terms of the number of occurrences in the ARG. Abbreviations: NSP, non-structural protein; ORF, open reading frame; S, Spike; S1-NTD, N-terminal domain in Subunit 1.}
\label{tab:rec_dels}
\begin{tabular}{rrcr}
\toprule
Start & Length & Region & Occurrences \\
\midrule
29781 & 1 & non-coding & 12754 \\
29762 & 1 & non-coding & 11383 \\
29779 & 1 & non-coding & 10951 \\
29769 & 1 & non-coding & 10809 \\
29774 & 1 & non-coding & 10735 \\
29700 & 1 & non-coding & 8562 \\
686 & 9 & ORF1ab / NSP1 & 6436 \\
29614 & 1 & ORF10 & 5039 \\
29555 & 1 & non-coding & 4814 \\
29543 & 1 & non-coding & 4804 \\
29541 & 1 & non-coding & 4802 \\
203 & 1 & non-coding & 3604 \\
222 & 1 & non-coding & 3442 \\
21991 & 3 & S / S1-NTD & 2088 \\
28254 & 1 & ORF8 & 1969 \\
28093 & 1 & ORF8 & 900 \\
21765 & 6 & S / S1-NTD & 482 \\
27532 & 1 & ORF7a & 404 \\
335 & 1 & ORF1ab / NSP1 & 370 \\
337 & 1 & ORF1ab / NSP1 & 346 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Parsimony comparison of sc2ts and UShER}
\label{sec:supplement_sc2ts_usher_parsimony}
We performed an indepth analysis of the sc2ts and UShER ARGs 
(we refer to the UShER tree as an ARG for simplicity here as a single 
tree is also an ARG) in terms of parsimony.
A direct comparison is difficult because of the different aligments 
used as input, the different sets of sites used for inference and 
the fact that we are including deletions for a subset of sites in
the sc2ts ARG.

>>>>

% See https://github.com/jeromekelleher/sc2ts-paper/issues/628
% Referenced by the main Results.
The ARG captures important deletion events (Table~\ref{tab:major_dels}), many of which are associated with the origins of major VOCs, for example, the deletions in ORF1ab and the N terminal domain of the S1 subunit associated with the Alpha lineage. and highly recurrent deletion events (Table~\ref{tab:rec_dels}).

\begin{table}
\centering
\caption{Major deletion events in the ARG. These events are inherited by at least 10,000 samples in the ARG. Count is the number of samples inheriting a deletion. The Pango lineage assigned to the node above which an event occurs is shown. Abbreviations: NSP, non-structural protein; ORF, open reading frame; S, Spike; S1-NTD, N terminal domain in S1 subunit.}
\label{tab:major_dels}
\begin{tabular}{rrcrcrr}
\toprule
Start & Length & Region & Node & Pango lineage & Count & Frequency (\%)\\
\midrule
28271 & 1 & non-coding & 1436808 & B.1 & 1118363 & 45.0561 \\
28248 & 6 & ORF8 & 220186 & B.1.617.2 (Delta) & 1114669 & 44.9073 \\
22029 & 6 & S / S1-NTD & 200039 & B.1.617.2 (Delta) & 1112504 & 44.8200 \\
11288 & 4 & ORF1ab / NSP6 & 1436802 & B.1.1.529 & 871782 & 35.1220 \\
28362 & 9 & N & 1436802 & B.1.1.529 & 871774 & 35.1216 \\
21633 & 9 & S / S1-NTD & 822854 & BA.2 (Omicron) & 532072 & 21.4359 \\
11292 & 5 & ORF1ab / NSP6 & 822854 & BA.2 (Omicron) & 531894 & 21.4287 \\
6513 & 3 & ORF1ab / NSP3 & 851246 & BA.1 (Omicron) & 340757 & 13.7283 \\
11283 & 5 & ORF1ab / NSP6 & 851246 & BA.1 (Omicron) & 339867 & 13.6924 \\
21765 & 6 & S / S1-NTD & 851246 & BA.1 (Omicron) & 339681 & 13.6849 \\
21987 & 9 & S / S1-NTD & 851246 & BA.1 (Omicron) & 339651 & 13.6837 \\
22194 & 3 & S / S1-NTD & 851246 & BA.1 (Omicron) & 312706 & 12.5982 \\
28271 & 1 & non-coding & 86456 & B.1.1.7 (Alpha) & 290446 & 11.7014 \\
11288 & 9 & ORF1ab / NSP6 & 86456 & B.1.1.7 (Alpha) & 290408 & 11.6998 \\
21991 & 3 & S / S1-NTD & 86456 & B.1.1.7 (Alpha) & 288831 & 11.6363 \\
21765 & 6 & S / S1-NTD & 86456 & B.1.1.7 (Alpha) & 290438 & 11.7010 \\
21765 & 6 & S / S1-NTD & 1265302 & BA.4 (Omicron) & 190425 & 7.6718 \\
686 & 9 & ORF1ab / NSP1 & 2698748 & BA.4 (Omicron) & 23330 & 0.9399 \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}
\centering
\caption{Highly recurrent deletions. These are the top 20 deletions in terms of the number of occurrences in the ARG. Abbreviations: NSP, non-structural protein; ORF, open reading frame; S, Spike; S1-NTD, N terminal domain in S1 subunit.}
\label{tab:rec_dels}
\begin{tabular}{rrcr}
\toprule
Start & Length & Region & Occurrences \\
\midrule
29781 & 1 & non-coding & 12754 \\
29762 & 1 & non-coding & 11383 \\
29779 & 1 & non-coding & 10951 \\
29769 & 1 & non-coding & 10809 \\
29774 & 1 & non-coding & 10735 \\
29700 & 1 & non-coding & 8562 \\
686 & 9 & ORF1ab / NSP1 & 6436 \\
29614 & 1 & ORF10 & 5039 \\
29555 & 1 & non-coding & 4814 \\
29543 & 1 & non-coding & 4804 \\
29541 & 1 & non-coding & 4802 \\
203 & 1 & non-coding & 3604 \\
222 & 1 & non-coding & 3442 \\
21991 & 3 & S / S1-NTD & 2088 \\
28254 & 1 & ORF8 & 1969 \\
28093 & 1 & ORF8 & 900 \\
21765 & 6 & S / S1-NTD & 482 \\
27532 & 1 & ORF7a & 404 \\
335 & 1 & ORF1ab / NSP1 & 370 \\
337 & 1 & ORF1ab / NSP1 & 346 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Parsimony comparison of sc2ts and UShER}
\label{sec:supplement_sc2ts_usher_parsimony}

We performed an indepth analysis of the sc2ts and UShER ARGs 
(we refer to the UShER tree as an ARG for simplicity here as a single 
tree is also an ARG) in terms of parsimony.
A direct comparison is difficult because of the different aligments 
used as input, the different sets of sites used for inference and 
the fact that we are including deletions for a subset of sites in
the sc2ts ARG.
